A linear inside-outside algorithm for correcting sequencing errors in structured RNA sequences

### 2012-10-07 Recomb/main_RECOMB.tex

\documentclass[11pt]{article}
\usepackage[top=1in, bottom=1in,right=1in,left=1in]{geometry}                % See geometry.pdf to learn the layout options. There are lots.
\geometry{letterpaper}                   % ... or a4paper or a5paper or ... 
%\geometry{landscape}                % Activate for for rotated page geometry
%\usepackage[parfill]{parskip}    % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm} %pour les newtheorem
\usepackage{epstopdf}
\usepackage{xspace}
\usepackage{relsize}
\usepackage[numbers]{natbib}
\usepackage[colorinlistoftodos]{todonotes}

@@

%\title{An inside-outside algorithms for mutant RNAs with applications to error detections in structured RNAs}
\title{A linear inside-outside algorithm for correcting sequencing errors in structured RNA sequences}
\author{Vladimir Reinharz$^{2}$, Yann Ponty$^{1}\footnote{To whom correspondence should be addressed}\;$, J\'er\^{o}me Waldisp\"{u}hl$^{2*}$}
\date{$^1$Laboratoire d'Informatique, \'Ecole Polytechnique, Palaiseau, 91120 France $^2$ School of Computer Science \& McGill Centre for Bioinformatics, McGill University, Montr\'eal, Canada H3C 2J7}

>>>>

\documentclass[11pt]{article}
\usepackage[top=1in, bottom=1in,right=1in,left=1in]{geometry}                % See geometry.pdf to learn the layout options. There are lots.
\geometry{letterpaper}                   % ... or a4paper or a5paper or ... 
%\geometry{landscape}                % Activate for for rotated page geometry
%\usepackage[parfill]{parskip}    % Activate to begin paragraphs with an empty line rather than an indent
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsthm} %pour les newtheorem
\usepackage{epstopdf}
\usepackage{xspace}
\usepackage{relsize}
\usepackage[numbers]{natbib}
\usepackage{natbibspacing}
\setlength{\bibspacing}{0em}
\usepackage[colorinlistoftodos]{todonotes}

@@

%\title{An inside-outside algorithms for mutant RNAs with applications to error detections in structured RNAs}
\title{A linear inside-outside algorithm for correcting sequencing errors in structured RNA sequences}
\author{Vladimir Reinharz$^{2}$, Yann Ponty$^{1}\footnote{To whom correspondence should be addressed}\;$, J\'er\^{o}me Waldisp\"{u}hl$^{2*}$}
\date{$^1$ CNRS/\'Ecole Polytechnique/Inria AMIB, LIX, Palaiseau, France \newline $^2$ School of Computer Science \& McGill Centre for Bioinformatics, McGill University, Montr\'eal, Canada H3C 2J7}

### 2012-10-07 Recomb/methods_RECOMB.tex

In principle, this computation should be performed $\Theta(m\cdot n^2)$ times to investigate any triplet $(m,i,j)$.
However, a close inspection of the recurrences reveal that the computation can be safely restricted to a subset of all intervals.
For instance, in the inside computation, such intervals $[i,j]$ do not break any base-pair, and their next position $j+1$ is either past the end of the sequence, or is base-paired prior to $i$. Similar constraints hold for the outside computation. Such constraints drastically limit the combinatorics of required computations, dropping from $\Theta(n^2)$ to $\Theta(n)$ the number of terms that need to be computed and stored. Consequently the overall complexity of the algorithm is $\Theta(n\cdot(|\Omega|+m^2))$ aritmetic operations and $\Theta(n\cdot(|\Omega|+m))$ memory.

>>>>

In principle, this computation should be performed $\Theta(m\cdot n^2)$ times to investigate any triplet $(m,i,j)$.
However, a close inspection of the recurrences reveals that the computation can be safely restricted to a subset of all intervals.
For instance, in the inside computation, such intervals $[i,j]$ are those that do not break any base-pair, and whose next position $j+1$ is either past the end of the sequence, or is base-paired prior to $i$. Similar constraints hold for the outside computation. Such constraints drastically limit the combinatorics of required computations, dropping from $\Theta(n^2)$ to $\Theta(n)$ the number of terms that need to be computed and stored. Consequently the overall complexity of the algorithm is $\Theta(n\cdot(|\Omega|+m^2))$ aritmetic operations and $\Theta(n\cdot(|\Omega|+m))$ memory.
